---
title: "petData"
author: "Susan Lively Klug"
date: "December 14, 2015"
output: html_document
---

###petData, a Shiny and Slidify assignment for Coursera Data Products class.

My submission for the assignment as a shiny application can be found at <http://slivelyklug.shinyapp.io/petData>.

#Purpose

##Viewing data about adoptable pets in animal shelters.

The data for this assignment is found through petfinder.com's API, they are
sent data about adoptable pets from animal shelters and rescue organizations
who wish to participate.  The website <http://www.petfinder.com> allows the 
user to search specific details about a desired future pet and displays 
contact information for adoption along with details and usually pictures of
matches.  

I am interested in how many overall adoptable pets there are, in a varity of 
locations and how many are male, female, how old they are.  I would like to 
know how long they have been in the shelter and the results for adoptions or otherwise.  However some of that information is not made readily available to 
the public.  Petfinder does allow some kinds of queries and for the purpose of
this assignment I chose to make a simple app that shows a subset of shelters
in the California area with interesting data.  To preserve the number of hits
using my key on petfinder I retrieved the data I wished to view and saved it as 
a data file.  This means the shiny app will not show current data, but a snapshot
of a single day.

Future expansions are already in the plans, the kinds of enhancements and 
possible interactivity is exciting.

#This Shiny application

##Working petData

petData 1.0:

* Upon launching shows an interactive widget on the right hand side
where the user may click on a radio button to choose a specific shelter.
* The top graph on the left side is static data of all available shelters in our
subset of the data.  
* The bottom graph on the left displays the types of animals
available at the shelter chosen by user, broken down into the reported sex of the
pet.  
* The shelter ID is also shown above the bottom graph confirming that the
data is for the chosen shelter.

petData future:  

* reactiveUI instead of static shelter ID list.
* Clicking on graphs to drill down to age, or showing age as a subset of sex.
* US map with data displaying based on hover or click on State.

petData tech:

To retrieve and clean the data, choose the shelters and find their names lots
of work was done on the command line.  This is poor form and will be resolved
in petData future editions wherein I can use my key to get live data and make
a reactiveUI.  In the meantime documentation about steps used to retrieve and
create the data is definitely wanted.  So here's the ugly brain dump truth.

```{r eval=FALSE}
xml.url <- "http://api.petfinder.com/pet.find?key=72609e7b0d2fac5982ee204316157c6d&id=CA803&count=1000&output=full&location=CA"
xmlfile <- xmlTreeParse(xml.url)
xmltop = xmlRoot(xmlfile)
```

Take the pets section.

```{r eval=FALSE}
pets <- xmltop[[3]]
petcatalog <- xmlSApply(pets, function(x) xmlSApply(x, xmlValue))
```

Put the data into a data-frame

```{r eval=FALSE}
pets_df <- data.frame(t(petcatalog),row.names=NULL)
unique(pets_df$shelterId)
```

Selected only the columns I care about.

```{r eval=FALSE}
df$size <- as.character(df$size)
df$size <- as.factor(df$size)
```

Using dplyr select.

```{r eval=FALSE}
mydata <- select(df, id:size)
```

Then lots of as.character, as.factors to convert the lists, find the better way asap.

```{r eval=FALSE}
mydata$sex <- as.character(mydata$sex)
mydata$sex <- as.factor(mydata$sex)
mydata$size <- as.character(mydata$size)
mydata$size <- as.factor(mydata$size)
mydata$age <- as.character(mydata$age)
mydata$age <- as.factor(mydata$age)
mydata$breeds <- as.character(mydata$breeds)
mydata$animal <- as.character(mydata$animal)
mydata$animal <- as.factor(mydata$animal)
mydata$name <- as.character(mydata$name)
mydata$shelterPetId <- as.character(mydata$shelterPetId)

class(mydata$id)
# [1] "numeric"
class(mydata$shelterId)
# [1] "list"

mydata$shelterId <- as.character(mydata$shelterId)
mydata$shelterId <- as.factor(mydata$shelterId)
```

Use test data set until app is running so I don't burn up my key.
Save data some different ways.

```{r eval=FALSE}
dump(mydata, "petsTestData.Rdmpd")
pets_df <- source("petsTestData.Rdmpd")
write.csv2(mydata, "pData.csv")

newdata <- read.csv2("pData.csv")
pets_df <- newdata
```

todo save and read in new data file....

todo reactiveUI - had it putting the shelterIds in the UI but the radio button
wasn't sending back the data to the server.

todo would like to have checkBox of shelters instead of just one.

todo if you run summary(mydata$shelterId) you will see many that have only 1 entry.

```{r eval=FALSE}
summary(mydata$shelterId)

# CA107 CA1192 CA1271 CA1280 CA1303 CA1365 CA1489 CA1522 CA1556 CA1612
# 20     23     18      1      2      3      8     12      2     54
# CA1626  CA166 CA1714 CA1794 CA1850 CA1910  CA193 CA2031 CA2162 CA2210
# 1      1     56      1     14      5     18      4      9      1
# CA2224 CA2260 CA2281 CA2329 CA2356 CA2382 CA2386 CA2395  CA250  CA434
# 20     12      1     11     15      2     33      2      1      8
# CA436  CA607  CA673  CA716  CA741  CA763  CA766  CA774  CA901  CA914
# 28     88     34      1      3    135    200     56     15      1
# CA990  MS113  NV160  WA526
# 78      1      1      1

bigS <- c("CA1192", "CA1612", "CA1714", "CA607",
          "CA673", "CA763", "CA774",
          "CA436", "CA2386", "CA990", "CA766")
mydata <- filter(pets_df, shelterId %in% bigS)
```
